Redundant calibration as a complex optimization problem

### 2017-01-24 paper.tex
fixed Jonathan's comments

\begin{abstract}
Traditionally redundant calibration is implemented via a hierarchical algorithmic strategy.  
We show that instead of splitting our redundant calibration problem into its phase-and-amplitude components,
which is the more traditional approach, we can treat the complex unknown model parameters and their complex conjugates as separate unknowns instead. 
We also derive the Levenberg-Maquardt and Gauss-Newton update steps associated with this alternative parameter-and-conjugate framework.
We focus more on the Levenberg-Maquardt algorithm in this paper as it exhibits better convergence properties when compared to Gauss-Newton.
Using the Levenberg-Maquardt algorithm, therefore, reduces the number of algorithmic layers that are needed to perform redundant calibration.
The Gauss-Newton algorithm is, however, still the de facto standard that is currently being used to perform redundant calibration.
We also compare two previously proposed methods to speedup redundant calibration, namely the alternating direction implicit method and the preconditioned conjugate gradient method. We found that the alternating 
direction implicit method outperforms the preconditioned conjugate gradient method, however the preconditioned conjugate gradient method is more robust as it is more straightforward to adapt to other calibration use cases. 
We make use of the parameter-and-conjugate framework in this paper, rather than the more traditional amplitude-and-phase approach as it makes it trivial to relate the alternating direction implicit method 
and the Levenberg-Marquardt algorithm with one another.
\end{abstract}

@@

\section{Introduction}
The propogation effects (i.e. the ionosphere, antenna gains, etc.) a celestial signal experiences along its propagation path is modelled with the radio interferometric measurement equation (RIME) \citep{ME1,RRIME1}.
Calibration is the procedure by which we remove the errors that these propagation effects induce onto our interferometric data and is normally accomplished using a 
least-squares solver. In the case where we only model the antenna gains, which is the most basic form of the RIME, we can define calibration as finding the antenna gains that minimizes the difference betweem our observed and predicted visibilities.

There are two recent developments in skymodel-based calibration which should be explicitly mentioned at this point, since they form the foundation this paper builds on. The first is the discovery of \textsc{StEfCal} (Statistically Efficient Calibration) \citep{Mitchell:MWA-cal,Salvini2014}, which 
is an alternating direction implicit (ADI) skymodel-based calibration algorithm. It works by assuming the gains are known while solving for their conjugates and then performing the reverse. 
This procedure results in the linearization of the calibration algorithm and reduces the computational complexity of calibration from $\mathbb{O}(N^3)$ to $\mathbb{O}(N^2)$, where $N$ denotes the number of antennas in the array. 
The second development is the idea: to not break the calibration problem 
into its real and imaginary parts and then to solve for the real and imaginary parts of the unknown parameters separately, but to rather treat the antenna gains and their 
conjugates as seperate parameters to estimate \citep{Smirnov2015}. \citet{Smirnov2015} also showed that the \textsc{StEfCal} algorithm itself is merely a special case of this
gain-and-conjugate framework.

Two baselines are redundant if their baseline difference vectors have the same length and orientation, which imply that they measure the same $uv$-modes. When we use a redundant array layout we dramatically 
reduce the number of unknowns. If our array is redundant enough we reduce the number of unknowns to such an extend that we are able to solve for the unknown 
observed visibilities themselves in addition to the antenna gains. Redundant calibration is the procedure by which we estimate the unknown observed visibilities in addition
to the antenna gains.

Redundant calibration was popularized in the 1980's. \citet{Noordam1982} used redundant calibration and the WSRT (Westerbork Synthesis Radio Telescope) to produce high-dynamic range images of 3C84.
The first implementations of redundant calibration made use of the logarithm function, because when the logarithm is applied to the scalar measurement equation, redundant 
calibration is transformed into a linear problem \citep{Wieringa1992,Camps2003,Liu2010}. The implementation associated with this approach is commonly referred to as \textsc{logcal}.
We briefly discuss \textsc{logcal} in Appendix~\ref{sec:logcal}. The next improvement in redundant calibration was to perform a firt-order taylor expansion, which is equivalent to the Gauss-Newton (GN) algorithm \citep{Kurien2016}, of the measurement equation and then to solve for the
perturbations of the model-parameters \citep{Liu2010}. \citet{Liu2010} perform the first order Taylor expansion of the measurement equation by splitting the unknown parameters into their phase-and-amplitude
components. The implementation of  this approach is called \textsc{lincal}. \textsc{lincal} is derived from first principals in Appendix~\ref{sec:lincal}. 
Currently, redundant calibration is usually implemented using a layered algorithmic strategy. \textsc{logcal} and \textsc{lincal} are both implemented in \textsc{omnical}\footnote{https://github.com/jeffzhen/omnical}, which is a layered redundant calibration pipeline \citep{Zheng2014,Ali2015}.
The solutions of \textsc{logcal} are used as initial input for \textsc{lincal}, since the GN algorithm requires that its starting parameter guesses be in the vicinity 
of the solution.

New instruments like LOFAR (Low Frequency Array) \citep{Noorishad2012}, PAPER (Precision Array for Probing the Epoch of Reionization) \citep{Ali2015} and HERA (Hydrogen Epoch of Reionization Array) \citep{deboer2015} have also recently sparked a renewed interest in redundant calibration. Using redundancy is especially advantageous when we are trying to detect the EoR (Epoch of Reionization) 21 cm signal,
which is inherently quite faint \citep{Parsons2012}. Since redundant baselines make statistically independent measurements of the same $uv$-modes redundancy can greatly 
improve the SNR (signal-to-noise ratio) of EoR detection experiments. This explains why PAPER and HERA, are highly redundant arrays, as they are both primarily used for EoR 21 cm detection experiments.

@@

Two ways of speeding up redundant calibration have also been proposed recently. The first technique makes use of the ADI method, in which we alternate between solving the gains, their conjugates and 
the true observed visibilities \citep{Wijnholds2012,Marthi2014}. Note that the fact that \citet{Marthi2014} is making use of the ADI method is not immediately clear
from the paper itself. We will elaborate on this later on in Section~\ref{sec:adi}. The second is to use the preconditioned conjugate gradient (PCG) method. PCG allows us to exploit the sparsity inherent in the redundant calibration problem, 
to speed it up \citep{Liu2010}. A proper comparison between these two approaches has not been done before.

In this paper we propose to use the complex framework introduced by \citet{Smirnov2015} instead of the traditional amplitude-and-phase approach used by \textsc{lincal}.
We also propose to use the Levenberg-Marquardt (LM) gradient-based algorithm instead of the Gauss-Newton (GN) algorithm as it exhibits better convergence properties.
Using the LM algorithm, reduces the number of algorithmic layers that are needed to perform redundant calibration. 
We also compare two recent redundant calibration speedup algorithms with each other, namely the ADI \citep{Marthi2014} and PCG \citep{Liu2010} methods. In this paper, we decided to mainly focus on the complex framework proposed by \citet{Smirnov2015} rather than the traditional phase-and-amplitude approach as it 
allows us to trivially prove that the ADI method proposed by \citep{Marthi2014} is merely a special case of the more general LM algorithm.

@@

We start the paper by presenting some useful preliminaries in Section~\ref{sec:preliminaries}. In Section~\ref{sec:p_statement}, we formulate redundant calibration 
as a least squares problem, i.e. Eq~\eqref{eq:least_squares_red}. In Section~\ref{sec:main_sec}, we use the complex framework proposed by \citet{Smirnov2015} to derive a numerical
algorithm (see Eq.~\eqref{eq:GN_update} to Eq.~\eqref{eq:update}), which we can use to solve Eq~\eqref{eq:least_squares_red}. In Section~\ref{sec:results}, we compare the computational complexity of two recently proposed redundant calibration speedup algorithms with one another.
We end the paper with some final conclusions.

@@

\subsubsection{Preconditioning}
\label{sec:precon}
Preconditioning is a technique which allows us to improve the condition number of a matrix. Let $\bM$ be a positive-definite Hermitian matrix.
We can now indirectly solve for $\bx$ by using 
\begin{equation}
\label{eq:prec}
\bM^{-1}\bA\bx = \bM^{-1}\bb.
\end{equation}
If $\bM$ is a good preconditioner then $\kappa(\bM^{-1}\bA) \ll \kappa(\bA)$, i.e. a good preconditioner lowers the condition number of a matrix which in turn increases the convergence speed 
of CG (see Eq.~\eqref{eq:cg_bound}). In this paper, we denote matrix inversion with $(*)^{-1}$. The best possible choice of $\bM$ would therefore be $\bA$, which would lead to $\kappa$ becoming unity (as the largest and smallest eigenvalue of the Identity matrix is equal to one).
If $\kappa$ is one then Eq.~\eqref{eq:cg_bound} simplifies and becomes $\mathbb{O}(m)$. However, if we knew the inverse of $\bA$ then obtaining $\bx$ would be trivial and there would be no need for the CG method in the first place. 
In short, a good preconditioner $\bM$ must approximate $\bA$ fairly accurately. However, if it is not much less expensive to 
compute the inverse of $\bM$ than that of $\bA$ we actually gain nothing (see Eq.~\eqref{eq:prec}).

If the diagonal entries of $\bA$ is much more significant than its off-diagonal entries then the 
Jacobian preconditioner is a good choice of $\bM$. The Jacobian preconditioner is computed as follows:
\begin{equation}
\bM = \bA\odot\bI, 
\end{equation}
where $\bI$ is the identity matrix and $\odot$ denotes the Hadamard product. Since the Jacobian preconditioner is a diagonal matrix its inverse can be obtained trivially.

@@

Recall that we mentioned that in general we can only solve Eq.~\eqref{eq:linear_system} with CG if $\bA$ is a square positive-definite Hermitian matrix. The CG method actually has an even broader applicability, it can be used to solve Eq.~\eqref{eq:linear_system} even if $\bA$ is an Hermitian positive semi-definite
square matrix as long as $\bb$ is in the column range of $\bA$ \citep{Lu2015}. In light of \citet{Lu2015}, the two observations in the numbered list we presented in the previous paragraph imply that Eq.~\eqref{eq:normal_equation}
can be solved by employing the CG method. The solution returned by CG is however only unique if $\bB^H\bB$ is positive-definite (invertible). The solution
obtained by CG, whether it be unique or not, does however minimize Eq.~\eqref{eq:norm_ls} in a least squares sense.

Note that if $\bA$ is positive semi-definite its smallest eigenvalue is zero, which implies that Eq.~\eqref{eq:kappa} is not defined. \citet{Lu2015} show that when $\bA$
is a square positive semi-definite Hermitian matrix we do not use the spectral condition number to bound the convergence complexity of CG, but the general spectral condition
number instead. The general spectral condition number is very similar to the spectral condition number; the only difference being that in Eq.~\eqref{eq:kappa} we replace  
$\lambda_{\textrm{min}}$ with $\lambda_{\textrm{min}}^+$, which denotes the smallest positive eigenvalue.

\subsection{Least Squares}
\label{sec:ls}
The method of nonlinear least squares is tanatamount to solving the following optimization problem:  
\begin{equation}
\label{eq:least_squares}
\min_{\bzeta} \Lambda(\bzeta) = \min_{\bzeta} \|\brho\|_F^2 = \min_{\bzeta} \|\bdelta - \bnu(\bzeta)\|_F^2, 
\end{equation}
where $\Lambda$ is known as the objective function, $\brho$ is the residual vector, $\bdelta$ is the observed data vector, $\bnu$ is the model-predicted data vector and $\bzeta$ is the model-parameter vector that minimizes Eq.~\eqref{eq:least_squares}.
In Eq.~\eqref{eq:least_squares}, we assume that the optimization problem is real valued.
The main objective of least squares is to find the model-parameters which minimizes the squared difference between the observed data and the model-predicted data.

The most standard way of solving Eq.~\eqref{eq:least_squares} is to use a gradient type of minimization algorithm. In this paper we will be focusing on the following gradient minimization algorithms: GN and LM \citep{Levenberg1944,Marquardt1963}. 
The GN update step is defined as
\begin{equation}
 \Delta \bzeta = (\bJ^T\bJ)^{-1} \bJ^T\brho,
\end{equation}
while the LM update takes the following form
\begin{equation}
 \Delta \bzeta = (\bJ^T\bJ+\lambda\bD)^{-1} \bJ^T\brho,
\end{equation}
where $\bJ$ is the Jacobian matrix (i.e. the derivative of the model $\bnu$ with respect to the parameter vector $\bzeta$), $\lambda$ denotes the 
damping factor, while $\bD$ is a diagonal matrix whose diagonal entries are equal to the diagonal entries of $\bJ^T\bJ$. In this paper, we denote 
matrix transposition with $(*)^T$.
These update steps can be used in an iterative manner to determine new estimates of $\bzeta$.

The gradient based minimization algorithms that are generally used to solve least squares problems (i.e. GN and LM) generally require the model $\bnu$ to be differentiable
towards each model parameter. When the least squares problem is complex it becomes less straightforward to apply these gradient based minimization methods. The reason being,
many complex functions are not analytic (no Taylor expansion around a fiducial point exists) and therefore not differentiable if the classic notion of differentiation is used. For instance if the 
classical definition of differentiation is used then $\frac{\partial \conj{z}}{\partial z}$, where $z \in \mathbb{C}$, does not exist.

The only way to circumvent the differentiability conundrum, associated with complex least square problems, is to recast our complex optimization problem in such a way that $\bnu$ becomes analytic in its argument.
The standard way of achieving this is to divide the complex optimization problem into its real and imaginary parts and then to solve for the real and imaginary parts of the parameters separately.

Recently, \citet{Sorber2012} proposed an alternative strategy, which involves the use of Wirtinger calculus \citep{Wirtinger1927}. The Wirtinger derivatives 
are defined to be 
\begin{align}
\label{eq:wir}
\frac{\partial}{\partial z} &= \frac{1}{2}\left ( \frac{\partial}{\partial x} -  i \frac{\partial}{\partial y} \right ),&\frac{\partial}{\partial \conj{z}} &= \frac{1}{2}\left ( \frac{\partial}{\partial x} +  i \frac{\partial}{\partial y} \right ). 
\end{align}
Using the above definitions we can now easily compute the following:
\begin{align}
\frac{\partial z}{\partial z} & = 1, & \frac{\partial \conj{z}}{\partial z}&=0, & \frac{\partial z}{\partial \conj{z}} & = 0, & \frac{\partial \conj{z}}{\partial \conj{z}}&=1.
\end{align}
In this alternative approach we treat the complex parameters and their conjugates as separate parameters.

\section{Problem Statement}
\label{sec:p_statement}
In Section~\ref{sec:s_mes}, we present the scalar measurement equation. Next, we define some indexing functions in Section~\ref{sec:indexing}. The measurement 
equation and the indexing functions from Section~\ref{sec:indexing} are then used in Section~\ref{sec:r_mes} to phrase redundant calibration as a least-squares problem, i.e. Eq.~\eqref{eq:least_squares_red}.
We present the Signal-to-noise ratio (SNR) definition we used in this paper in Section~\ref{sec:snr}. The limitations of this paper are discussed in Section~\ref{sec:scope}.

@@

The real and imaginary components of the thermal noise is normally distributed with a mean of zero and a standard deviation proportional to   
\begin{equation}
\sigma \propto \frac{T_{\textrm{sys}}}{\sqrt{\Delta \nu \tau}}, 
\end{equation}
where $T_{\textrm{sys}}$ is equal to the system temperature, $\Delta \nu$ is the observational bandwidth and $\tau$ is the integration time per visibility.

\subsection{Indexing Functions}
\label{sec:indexing}
In general there are two ways of indexing the true visibilities associated with different baselines. We can use the composite index $pq$, which is the strategy that is followed in Eq.~\eqref{eq:vis_definition},
or we can use a single unique baseline index which we assign in numerical order to each composite index; assuming the composite indexes are arranged in standard baseline order.
The composite indexes $pq$ are in standard baseline order when they are arranged as follow: $12, 13, 1N, 23\cdots$. Let 
\begin{equation}
\alpha_{pq} =
\begin{cases}
(q-p) + (p-1)\left (N-\frac{1}{2}p \right ) & \textrm{if}~p<q\\
%B + (p-q) + (q-1)\left (N-\frac{1}{2}q \right )) & \textrm{if}~p>q\\
%B+q + \frac{1}{2}(p-1)(p-2) & \textrm{if}~p>q\\
0 & \textrm{otherwise}
\end{cases}.
\end{equation}
We can use $\alpha_{pq}$ to map composite indexes to unique baseline indexes, i.e. 
\begin{equation}
12,13,\cdots,1N,23,\cdots \xrightarrow{\alpha_{pq}} 1,2,3,\cdots,B.
\end{equation}
We denote the number of baselines in the array with $B$, i.e. $B = \frac{N^2-N}{2}$.

>>>>

\begin{abstract}
In this paper we show that instead of splitting our redundant calibration problem into its phase-and-amplitude components (the more traditional approach) we can treat the complex unknown model parameters and their complex conjugates as separate unknowns. 
We also derive the Levenberg-Maquardt and Gauss-Newton update steps associated with this alternative parameter-and-conjugate framework.
We focus more on the Levenberg-Maquardt algorithm in this paper as it exhibits better convergence properties when compared to Gauss-Newton.
Traditionally redundant calibration is implemented via a hierarchical algorithmic strategy. Using the Levenberg-Maquardt algorithm, therefore, reduces the number of algorithmic layers that are needed to perform redundant calibration.
The Gauss-Newton algorithm is, however, still the de facto standard that is currently being used to perform redundant calibration.
We also compare two previously proposed methods to accelerate redundant calibration, namely the alternating direction implicit method and the preconditioned conjugate gradient method. We found that the alternating 
direction implicit method outperforms the preconditioned conjugate gradient method. However, the preconditioned conjugate gradient method is more robust as it is more straightforward to adapt to other calibration use cases. 
We make use of the parameter-and-conjugate framework in this paper, rather than the more traditional amplitude-and-phase approach as it makes it trivial to relate the alternating direction implicit method 
and the Levenberg-Marquardt algorithm with one another.
\end{abstract}

@@

\section{Introduction}
The effects (i.e. the ionosphere, antenna gains, etc.) a celestial signal experiences along its propagation path are modelled with the radio interferometric measurement equation (RIME) \citep{ME1,RRIME1}.
Calibration is the procedure by which we remove the errors that these propagation effects introduce into interferometric data and is normally accomplished using a 
least-squares solver. In the case where we only model the antenna gains, which is the most basic form of the RIME, we can define calibration as finding the antenna gains that minimize the difference between our observed and predicted visibilities.

There are two recent developments in skymodel-based calibration which should be explicitly mentioned at this point, since they form the foundation this paper builds on. The first is the discovery of \textsc{StEfCal} (Statistically Efficient Calibration) \citep{Mitchell:MWA-cal,Salvini2014}, which 
is an alternating direction implicit (ADI) skymodel-based calibration algorithm. It works by assuming the gains are known while solving for their conjugates and then performing the reverse. 
This procedure results in the linearization of the calibration algorithm and reduces the computational complexity of calibration from $\mathbb{O}(N^3)$ to $\mathbb{O}(N^2)$, where $N$ denotes the number of antennas in the array. 
The second development is the idea: instead of breaking the calibration problem 
into its real and imaginary parts and then solving for the real and imaginary parts of the unknown parameters separately we can treat the antenna gains and their 
conjugates as separate parameters to estimate \citep{Smirnov2015}. \citet{Smirnov2015} also showed that the \textsc{StEfCal} algorithm itself is merely a special case of this
gain-and-conjugate framework.

Two baselines are redundant if their baseline difference vectors have the same length and orientation, which imply that they measure the same $uv$-modes. When we use a redundant array layout, we dramatically 
reduce the number of unknowns. If our array is redundant enough we reduce the number of unknowns to such an extent that we are able to solve for the unknown 
observed visibilities themselves in addition to the antenna gains. Redundant calibration is the procedure by which we estimate both the unknown observed visibilities and the antenna gains.

Redundant calibration was popularized in the 1980's. \citet{Noordam1982} used redundant calibration and the WSRT (Westerbork Synthesis Radio Telescope) to produce high-dynamic range images of 3C84.
The first implementations of redundant calibration made use of the logarithm function, because when the logarithm is applied to the scalar measurement equation, redundant 
calibration is transformed into a linear problem \citep{Wieringa1992,Camps2003,Liu2010}. The implementation associated with this approach is commonly referred to as \textsc{logcal}.
We briefly discuss \textsc{logcal} in Appendix~\ref{sec:logcal}. The next improvement in redundant calibration was to perform a first order Taylor expansion, which is equivalent to the Gauss-Newton (GN) algorithm \citep{Kurien2016}, of the measurement equation and then to solve for the
perturbations of the model-parameters \citep{Liu2010}. \citet{Liu2010} perform the first order Taylor expansion of the measurement equation by splitting the unknown parameters into their phase-and-amplitude
components. The implementation of  this approach is called \textsc{lincal}. \textsc{lincal} is derived from first principals in Appendix~\ref{sec:lincal}. 
Currently, redundant calibration is usually implemented using a layered algorithmic strategy. \textsc{logcal} and \textsc{lincal} are both implemented in \textsc{omnical}\footnote{https://github.com/jeffzhen/omnical}, which is a layered redundant calibration pipeline \citep{Zheng2014,Ali2015}.
The solutions of \textsc{logcal} are used as initial input for \textsc{lincal}, since the GN algorithm requires that its starting parameter guesses be in the vicinity 
of the true solution.

New instruments like LOFAR (Low Frequency Array) \citep{Noorishad2012}, PAPER (Precision Array for Probing the Epoch of Reionization) \citep{Ali2015} and HERA (Hydrogen Epoch of Reionization Array) \citep{deboer2015} have also sparked a renewed interest in redundant calibration. Using redundancy is especially advantageous when we are trying to detect the EoR (Epoch of Reionization) 21 cm signal,
which is inherently quite faint \citep{Parsons2012}. Since redundant baselines make statistically independent measurements of the same $uv$-modes redundancy can greatly 
improve the SNR (signal-to-noise ratio) of EoR detection experiments. This explains why PAPER and HERA are highly redundant arrays, as they are both primarily used for EoR 21 cm detection experiments.

@@

Two ways of speeding up redundant calibration have recently been proposed. The first technique makes use of the ADI method, in which we alternate between solving the gains, their conjugates and 
the true observed visibilities \citep{Wijnholds2012,Marthi2014}. Note that the fact that \citet{Marthi2014} make use of the ADI method is not immediately clear
from the paper itself. We will elaborate on this in Section~\ref{sec:adi}. The second is to use the preconditioned conjugate gradient (PCG) method. PCG allows us to exploit the sparsity inherent in the redundant calibration problem 
to speed it up \citep{Liu2010}. A proper comparison between these two approaches has not been done before.

It is also interesting to mention here that a completely different calibration strategy can also be used to calibrate redundant arrays, \textsc{CorrCal} \citep{Sievers2017}\footnote{https://github.com/sievers/corrcal2}. In \textsc{CorrCal} the unknown 
visibilities are no longer solved for. The sky is actually modelled, via two covariance matrices. 
The first covariance matrix is associated with known point sources. The second covariance matrix models the remaining sky in a statistical sense. The unknown gains, however, are still 
estimated. We do not elaborate any further on this approach in this paper.

In this paper we propose to use the complex framework introduced by \citet{Smirnov2015} instead of the traditional amplitude-and-phase approach used by \textsc{lincal}.
We also propose to use the Levenberg-Marquardt (LM) gradient-based algorithm instead of the Gauss-Newton (GN) algorithm as it exhibits better convergence properties.
Using the LM algorithm reduces the number of algorithmic layers that are needed to perform redundant calibration. 
We also compare two recent redundant calibration speedup algorithms with each other, namely the ADI \citep{Marthi2014} and PCG \citep{Liu2010} methods. In this paper, we decided to focus on the complex framework proposed by \citet{Smirnov2015} rather than the traditional phase-and-amplitude approach as it 
allows us to trivially prove that the ADI method proposed by \citep{Marthi2014} is merely a special case of the more general LM algorithm.

@@

We start the paper by presenting some useful preliminaries in Section~\ref{sec:preliminaries}. In Section~\ref{sec:p_statement}, we formulate redundant calibration 
as a least squares problem, i.e. Eq~\eqref{eq:least_squares_red}. In Section~\ref{sec:main_sec} we use the complex framework proposed by \citet{Smirnov2015} to derive a numerical
algorithm (see Eq.~\eqref{eq:GN_update} to Eq.~\eqref{eq:update}), which we can use to solve Eq~\eqref{eq:least_squares_red}. In Section~\ref{sec:results}, we compare the computational complexity of two recently proposed redundant calibration speedup algorithms with one another.
We end the paper with some final conclusions.

@@

\subsubsection{Preconditioning}
\label{sec:precon}
Preconditioning is a technique which allows us to improve the condition number of a matrix. Let $\bM$ be a positive-definite Hermitian matrix.
We can now indirectly solve for $\bx$ by using 
\begin{equation}
\label{eq:prec}
\bM^{-1}\bA\bx = \bM^{-1}\bb.
\end{equation}
If $\bM$ is a good preconditioner then $\kappa(\bM^{-1}\bA) \ll \kappa(\bA)$, i.e. a good preconditioner lowers the condition number of a matrix which in turn increases the convergence speed 
of CG (see Eq.~\eqref{eq:cg_bound}). In this paper, we denote matrix inversion with $(*)^{-1}$. The best possible choice of $\bM$ would therefore be $\bA$, which would lead to $\kappa$ becoming unity (as the largest and smallest eigenvalue of the Identity matrix are equal to one).
If $\kappa$ is one then Eq.~\eqref{eq:cg_bound} simplifies and becomes $\mathbb{O}(m)$. However, if we knew the inverse of $\bA$ then obtaining $\bx$ would be trivial and there would be no need for the CG method in the first place. 
In short, a good preconditioner $\bM$ must approximate $\bA$ fairly accurately. However, if it is not much less expensive to 
compute the inverse of $\bM$ than that of $\bA$ we actually gain nothing (see Eq.~\eqref{eq:prec}).

If the diagonal entries of $\bA$ are much more significant than its off-diagonal entries then the 
Jacobian preconditioner is a good choice of $\bM$. The Jacobian preconditioner is computed as follows:
\begin{equation}
\bM = \bA\odot\bI, 
\end{equation}
where $\bI$ is the identity matrix and $\odot$ denotes the Hadamard product. Since the Jacobian preconditioner is a diagonal matrix its inverse can be obtained trivially.

@@

Recall that we mentioned that in general we can only solve Eq.~\eqref{eq:linear_system} with CG if $\bA$ is a square positive-definite Hermitian matrix. The CG method actually has an even broader applicability: it can be used to solve Eq.~\eqref{eq:linear_system} even if $\bA$ is a Hermitian positive semi-definite
square matrix as long as $\bb$ is in the column range of $\bA$ \citep{Lu2015}. In light of \citet{Lu2015}, the two observations in the numbered list we presented in the previous paragraph imply that Eq.~\eqref{eq:normal_equation}
can be solved by employing the CG method. The solution returned by CG is, however, only unique if $\bB^H\bB$ is positive-definite (invertible). The solution
obtained by CG, whether it is unique or not, does minimize Eq.~\eqref{eq:norm_ls} in a least squares sense.

Note that if $\bA$ is positive semi-definite, its smallest eigenvalue is zero. This implies that Eq.~\eqref{eq:kappa} is not defined. \citet{Lu2015} show that when $\bA$
is a square positive semi-definite Hermitian matrix we do not use the spectral condition number to bound the convergence complexity of CG, but the general spectral condition
number instead. The general spectral condition number is very similar to the spectral condition number; the only difference being that in Eq.~\eqref{eq:kappa} we replace  
$\lambda_{\textrm{min}}$ with $\lambda_{\textrm{min}}^+$, which denotes the smallest positive eigenvalue.

\subsection{Least Squares}
\label{sec:ls}
The method of nonlinear least squares is tanatamount to solving the following optimization problem:  
\begin{equation}
\label{eq:least_squares}
\min_{\bzeta} \Lambda(\bzeta) = \min_{\bzeta} \|\brho\|_F^2 = \min_{\bzeta} \|\bdelta - \bnu(\bzeta)\|_F^2, 
\end{equation}
where $\Lambda$ is known as the objective function, $\brho$ is the residual vector, $\bdelta$ is the observed data vector, $\bnu$ is the model-predicted data vector and $\bzeta$ is the model-parameter vector that minimizes Eq.~\eqref{eq:least_squares}.
In Eq.~\eqref{eq:least_squares}, we assume that the optimization problem is real valued.
The main objective of least squares is to find the model-parameters which minimize the squared difference between the observed data and the model-predicted data.

The most standard way of solving Eq.~\eqref{eq:least_squares} is to use a gradient-type minimization algorithm. In this paper we will be focusing on the following gradient minimization algorithms: GN and LM \citep{Levenberg1944,Marquardt1963}. 
The GN update step is defined as
\begin{equation}
 \Delta \bzeta = (\bJ^T\bJ)^{-1} \bJ^T\brho,
\end{equation}
while the LM update takes the following form
\begin{equation}
 \Delta \bzeta = (\bJ^T\bJ+\lambda\bD)^{-1} \bJ^T\brho,
\end{equation}
where $\bJ$ is the Jacobian matrix (i.e. the derivative of the model $\bnu$ with respect to the parameter vector $\bzeta$), $\lambda$ denotes the 
damping factor and $\bD$ is a diagonal matrix whose diagonal entries are equal to the diagonal entries of $\bJ^T\bJ$. In this paper, we denote 
matrix transposition with $(*)^T$.
These update steps can be used in an iterative manner to determine new estimates of $\bzeta$.

The gradient based minimization algorithms that are generally used to solve least squares problems (i.e. GN and LM) generally require the model $\bnu$ to be differentiable
towards each model parameter. When the least squares problem is complex it becomes less straightforward to apply these gradient based minimization methods, as
many complex functions are not analytic (no Taylor expansion around a fiducial point exists) and therefore not differentiable if the classic notion of differentiation is used. For instance, if the 
classical definition of differentiation is used then $\frac{\partial \conj{z}}{\partial z}$, where $z \in \mathbb{C}$, does not exist.

The only way to circumvent the differentiability conundrum associated with complex least square problems is to recast our complex optimization problem in such a way that $\bnu$ becomes analytic in its argument.
The standard way of achieving this is to divide the complex optimization problem into its real and imaginary parts and then to solve for the real and imaginary parts of the parameters separately.

Recently, \citet{Sorber2012} proposed an alternative strategy, which involves the use of Wirtinger calculus \citep{Wirtinger1927}. The Wirtinger derivatives 
are defined to be 
\begin{align}
\label{eq:wir}
\frac{\partial}{\partial z} &= \frac{1}{2}\left ( \frac{\partial}{\partial x} -  i \frac{\partial}{\partial y} \right ),&\frac{\partial}{\partial \conj{z}} &= \frac{1}{2}\left ( \frac{\partial}{\partial x} +  i \frac{\partial}{\partial y} \right ). 
\end{align}
Using the above definitions we can now easily compute the following:
\begin{align}
\frac{\partial z}{\partial z} & = 1, & \frac{\partial \conj{z}}{\partial z}&=0, & \frac{\partial z}{\partial \conj{z}} & = 0, & \frac{\partial \conj{z}}{\partial \conj{z}}&=1.
\end{align}
In this alternative approach we treat the complex parameters and their conjugates as separate variables.

\section{Problem Statement}
\label{sec:p_statement}
In Section~\ref{sec:s_mes}, we present the scalar measurement equation. Next, we define some indexing functions in Section~\ref{sec:indexing}. The measurement 
equation and the indexing functions from Section~\ref{sec:indexing} are then used in Section~\ref{sec:r_mes} to phrase redundant calibration as a least-squares problem, i.e. Eq.~\eqref{eq:least_squares_red}.
We present the signal-to-noise ratio (SNR) definition we used in this paper in Section~\ref{sec:snr}. The limitations of this paper are discussed in Section~\ref{sec:scope}.

@@

The real and imaginary components of the thermal noise are normally distributed with a mean of zero and a standard deviation $\sigma$, where   
\begin{equation}
\sigma \propto \frac{T_{\textrm{sys}}}{\sqrt{\Delta \nu \tau}}. 
\end{equation}
Here $T_{\textrm{sys}}$ is equal to the system temperature, $\Delta \nu$ is the observational bandwidth and $\tau$ is the integration time per visibility.

\subsection{Indexing Functions}
\label{sec:indexing}
In general there are two ways of indexing the true visibilities associated with different baselines. We can use the composite index $pq$, which is the strategy that is followed in Eq.~\eqref{eq:vis_definition},
or we can use a single unique baseline index which we assign in numerical order to each composite index, assuming the composite indexes are arranged in standard baseline order.
The composite indexes $pq$ are in standard baseline order when they are arranged as follow: $12, 13, 1N, 23\cdots$. Let 
\begin{equation}
\alpha_{pq} =
\begin{cases}
(q-p) + (p-1)\left (N-\frac{1}{2}p \right ) & \textrm{if}~p<q\\
%B + (p-q) + (q-1)\left (N-\frac{1}{2}q \right )) & \textrm{if}~p>q\\
%B+q + \frac{1}{2}(p-1)(p-2) & \textrm{if}~p>q\\
0 & \textrm{otherwise}
\end{cases}.
\end{equation}
We can use $\alpha_{pq}$ to map composite indexes to unique baseline indexes, i.e. 
\begin{equation}
12,13,\cdots,1N,23,\cdots \xrightarrow{\alpha_{pq}} 1,2,3,\cdots,B.
\end{equation}
We denote the number of baselines in the array with $B$, i.e. $B = \frac{N^2-N}{2}$.

Recall that redundant baselines sample the exact same visibilities in the $uv$-plane, i.e. if baseline $pq$ and $rs$ are redundant then $y_{pq} = y_{rs}$. 
Let us now group all the baselines that measure the same $uv$-modes in a redundant array into the same redundant baseline group. Assigning the baselines of a redundant array to different redundant baseline groups has the advantage that it reduces the number of single index values we require. We now only require single index values for each redundant baseline group instead of a unique single index label for 
each baseline. Let $\phi_{pq}$ be the function that maps composite antenna pairs to the appropriate single redundant baseline group index. Note that $\phi_{pq}$ is not symmetric as 
$\phi_{pq} = 0$ if $p>q$. We can also define the following symmetric variant of $\phi_{pq}$:
\begin{equation}
\zeta_{pq} = 
\begin{cases}
\phi_{pq}~\textrm{if}~p \leq q\\
\phi_{qp}~\textrm{if}~p>q
\end{cases}.
\end{equation} 
It is possible to construct a simple analytic expression for $\zeta_{pq}$ when our array is in an east-west regular configuration, namely $\zeta_{pq} = |q-p|$. 
It becomes increasingly difficult to construct analytic expressions of $\zeta_{pq}$ for more complicated array layouts. 
We will refer to $\zeta_{pq}$ as the symmetric geometric function. The empirically constructed symmetric geometry
functions of three different redundant layouts are depicted in Fig.~\ref{fig:geometry_function}. We denote the range of $\zeta_{pq}$ with $\mathcal{R}(\zeta_{pq})$. The maximal element 
that $\zeta_{pq}$ can ascertain is denoted by $L$ and can be interpreted as the maximal number of unique redundant baseline groups which can be formed for a given 
array layout. Alternatively, we can interpret the symmetric geometry functions in Fig.~\ref{fig:geometry_function} as geometry matrices. In this alternative paradigm, $\zeta_{pq}$ denotes a matrix entry. The dimension of these geometry 
matrices are $N\times N$.
